Analytic Doppler Imaging

### 2019-08-23 tex/ms.tex
paper coming along

\begin{abstract} 
We derive a closed form, analytic solution to the problem of Doppler imaging
of stellar surfaces in the limit of negligible differential rotation and
convective blueshift. The model for the observed spectrum is linear in 
both the unbroadened spectrum and the surface map of the star, provided the
latter is expressed as a sum of spherical harmonics. As a result, the 
posterior distributions of both the unbroadened spectrum and the stellar map
conditioned on each other have a closed form, analytic solution that is 
computationally trivial to evaluate. When neither the unbroadened spectrum
nor the stellar map is known a priori, the global optimum can usually be
found via a fast iterative approach that also yields a lower bound on the
uncertainties.
\xxx{%
Data-driven modeling at its finest $\cdot$ %
Works on blended lines $\cdot$ %
Uses the whole spectrum $\cdot$ %
Uncertainties FTW $\cdot$ %
Probably works on slow rotators $\cdot$ %
Differentiable! $\cdot$ %
Fast! $\cdot$ %
Use in MCMC to marginalize over inclination and stuff $\cdot$ %
}
%
\href{https://github.com/rodluger/paparazzi}{\color{linkcolor}\faGithub}
\end{abstract}

%
\section{Introduction}
%
\xxx{Check out \citet{Luger2019} and \citet{Bedell2019} and stuff.}

@@

\subsection{The full solution}
In the previous two sections we showed how, if the stellar
rest-frame spectrum is spatially constant and known \emph{a priori} 
(from, say, a template spectrum), the posterior distribution
for the surface map $\bvec{y}$ is \emph{analytic} 
(Equations~\ref{eq:inverse:y1hat}
and \ref{eq:inverse:y1cov}). Conversely, if the surface map is known,
the posterior for the spectrum is also analytic 
(Equations~\ref{eq:inverse:shat} and \ref{eq:inverse:scov}). Both sets of
solutions require Gaussian priors on the quantities of interest, but these
are easily interpretable. A Gaussian prior on
the spherical harmonic coefficients of the surface map corresponds 
to the angular power spectrum of the surface features, whereas a Gaussian
prior on the spectrum can be used to correctly propagate uncertainties in
the stellar spectrum from observational uncertainties on the stellar
metallicity, temperature, etc.

@@

\subsection{Limb darkening}
\label{sec:ld}
Thus far we have avoided mention of limb darkening, whose effect
on the observed spectrum can often be significant. A proper treatment of
limb darkening requires solution of radiative transfer equations and
should account for the temperature and composition gradients in the
stellar atmosphere. However, in the spirit of data-driven inference, we 
can approximate the effect of limb darkening by weighting the spatial
component of our map by the limb darkening profile of the star, which can 
either be imposed or learned from the data. For a polynomial limb darkening
law of the form
%
\begin{align}
    \label{eq:ld:I}
    \frac{I(\mu)}{I(\mu = 1)} = 1 - \sum_{n=1}^{n_\mathrm{max}} u_n(1 - \mu)^n
    \quad,
\end{align}
%
where $\mu = z = \sqrt{1 - x^2 - y^2}$ is the radial coordinate on the 
projected disk and $u_n$ is a limb darkening coefficient, the effect of 
limb darkening on the stellar map can be expressed exactly as a linear 
operation on the spherical harmonic coefficient vector 
\citep{Luger2019,Agol2019}. This can be understood by noting that all terms
in Equation~(\ref{eq:ld:I}) are strictly polynomials in $x$, $y$, and $z$,
all of which can be expressed exactly as sums of spherical harmonics
\citep{Luger2019}. When weighting the surface intensity by the limb darkening
profile, the resulting intensity is simply a product of spherical harmonics,
which is itself a sum of spherical harmonics. Given a limb darkening law
of degree $n_\mathrm{max}$, we can always construct a matrix $\bvec{L}$
that transforms a spherical harmonic vector $\bvec{y}$ of degree
$l_\mathrm{max}$ to a limb-darkened spherical harmonic vector $\bvec{y'}$
of degree $l_\mathrm{max} + n_\mathrm{max}$. As an example, consider 
a map of degree $l_\mathrm{max} = 1$ and the linear limb darkening law
($n_\mathrm{max} = 1$).
The transformation matrix from $\bvec{y}$ to $\bvec{y'}$ is
%
\begin{align}
    \label{eq:ld:L}
    \setstackgap{L}{1.5\baselineskip}
    \bvec{L} =
        \frac{1}{1 - \frac{u_1}{3}}
        \begin{pmatrix}
        \quadquad1-u_1\quadquad & 0                         & \frac{u_1}{\sqrt{3}}      & 0                         \\
        0                       & \quadquad1-u_1\quadquad   & 0                         & 0                         \\
        \frac{u_1}{\sqrt{3}}    & 0                         & \quadquad1-u_1\quadquad   & 0                         \\
        0                       & 0                         & 0                         & \quadquad1-u_1\quadquad   \\
        0                       & 0                         & 0                         & 0                         \\
        0                       & \frac{u_1}{\sqrt{5}}      & 0                         & 0                         \\
        0                       & 0                         & \frac{2u_1}{\sqrt{15}}    & 0                         \\
        0                       & 0                         & 0                         & \frac{u_1}{\sqrt{5}}      \\
        0                       & 0                         & 0                         & 0          
        \end{pmatrix}
        \quad.
\end{align}
%
The columns of $\bvec{L}$ are constructed from the coefficient vectors of
each transformed spherical harmonic, which are in turn computed by 
multiplying each spherical harmonic by the spherical harmonic decomposition 
of the particular limb darkening law.
Figure~\ref{fig:ld} shows this transformation with $u_1 = 1$ applied to each
of the first four spherical harmonics.
%
\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=0.6\linewidth]{figures/ld.pdf}
    \oscaption{ld}{%
        Illustration of the limb darkening transformation given by
        Equation~(\ref{eq:ld:L}) for a linear limb darkening law with
        $u_1 = 1$. Each row shows a spherical harmonic before
        (left) and after (right) transformation by $\bvec{L}$.
        \label{fig:ld}
    }
    \end{centering}
\end{figure}
%

@@

The model remains linear in both the spherical harmonic coefficients and the
spectral components, and the procedures outlined in \S\ref{sec:inverse} may
still be followed to solve the inverse problem, provided we unroll these
matrices into vectors, letting
$\mathbf{s} = \mathrm{vec}(\boldsymbol{\mathbb{S}})$
and
$\mathbf{y} = \mathrm{vec}(\boldsymbol{\mathbb{Y}})$.
A detailed discussion of this general case is deferred to future work,
but we present an example of a two-component stellar spectrum in the
following section.

@@

\xxx{note on the intensities at A and B.}

Given $\bvec{y_1}$ and $\bvec{s}$, we use Equation~(\ref{eq:deconv:F}) to
compute the data vector $\bvec{f}$
at $M = 16$ equally spaced rotational phases; these phases are shown in orthographic
projection (as an observer would see the star) in the center
panel of Figure~\ref{fig:spot_setup}. The synthetic spectra for each phase
are shown as the orange curves in the bottom panel of the figure (note the
different scaling, as all spectral lines become shallower due to the
broadening). We assume a wavelength-independent,
quadratic limb darkening law with $u_1 = 0.5$ and $u_2 = 0.25$. Finally,
we divide the flux by the baseline at each epoch (Equation~\ref{eq:norm:b})
and add random Gaussian noise with amplitude $\sigma_\bvec{f} = 10^{-4}$ to
simulate a very high signal-to-noise ratio (SNR) spectrum. The mean
standard deviation in each wavelength bin across all 16 epochs in the
noiseless data vector is about $10^{-2}$, so our choice of $\sigma_\bvec{f}$
corresponds to SNR $\sim$ 100 in each of the 201 wavelength bins of $\bvec{f}$.
%
%For reference, \citet{Vogt1987} report SNRs per 60 m\AA\, pixel, while our
%resolution elements are approximately $\lambda_r\Delta\lnlam/K \approx 100$ m\AA\,
%across, so our SNR in the same units as \citet{Vogt1987} is about 80.
%
% TODO: Perhaps Vogt computes the "signal" as the line depth, rather than
% as the amplitude of the perturbations? Check this.

@@

\subsection{Known baseline and spectrum}
%
In our first experiment, we attempt to infer the stellar map assuming
we have perfect knowledge of the rest frame spectrum $\bvec{s}$ and
the baseline $\bvec{b}$ (Equation~\ref{eq:norm:b}). 
This is analogous to the case where a good physical model exists for
the stellar spectrum and simultaneous photometry is available to
constrain the baseline variations.
As we showed in
\S\ref{sec:solve_y}, the posterior distribution for the map coefficients 
$\bvec{y_1}$ is analytic in this case, with mean given by Equation~(\ref{eq:inverse:y1hat})
and covariance given by Equation~(\ref{eq:inverse:y1cov}). Note that
we must first scale the data vector $\bvec{f}$ by $\bvec{b}$ (which in this
case is known).

@@

\subsection{Known spectrum, unknown baseline}
%
We next investigate the case where the spectrum is known exactly, but
the baseline is not.

@@

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=0.49\linewidth]{figures/spot_y1bs_timeseries.pdf}
    \includegraphics[width=0.49\linewidth]{figures/spot_y1bs_rect.pdf}
    \\[0.5em]
    \includegraphics[width=0.405\linewidth]{figures/spot_y1bs_baseline.pdf}
    \includegraphics[width=0.575\linewidth]{figures/spot_y1bs_coeffs.pdf}
    \\[0.5em]
    \includegraphics[width=0.98\linewidth]{figures/spot_y1bs_spectrum.pdf}
    \oscaption{spot_y1bs}{%
        Similar to Figure~\ref{fig:spot_y1}, but assuming we have no
        information about either the spectrum, the baseline, or the map
        coefficients. The bottommost panel now shows the true rest frame
        spectrum (blue line), the initial guess based on a deconvolution
        (dotted orange line), and the inferred spectrum (orange line, with
        shaded $1\sigma$ uncertainty).
        \label{fig:spot_y1bs}
    }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=0.98\linewidth]{figures/inclinations.pdf}
    \oscaption{inclinations}{%
        Recovered \spot map as a function of stellar inclination. The
        rest frame spectrum is assumed to be known exactly, but the
        baseline is not. At very low inclinations, the small
        projected velocities begin to degrade the spatial resolution of
        the inferred map, while at inclinations close to $90^\circ$ there
        is a degeneracy between the northern and southern hemispheres.
        \label{fig:inclinations}
    }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=0.98\linewidth]{figures/snr.pdf}
    \oscaption{snr}{%
        Recovered \spot map as a function of SNR per resolution element.
        The quality of the inferred map in the northern hemisphere improves
        with SNR, but at very high SNR the region surrounding the global
        optimum becomes so small that our optimization methods struggle to
        converge on the correct solution.
        \label{fig:snr}
    }
    \end{centering}
\end{figure}

\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/twospec.pdf}
    \oscaption{twospec}{%
        An example of a star whose spectrum is different within the spot
        (\texttt{B}) and outside of the spot (\texttt{A}). Everywhere on 
        the surface, the spectrum is a linear combination of two
        ``eigenspectra'' ($s_1$ and $s_2$).
        As in Figure~\ref{fig:spot_y1bs}, we assume
        we do not know the map, the baseline, or either of the eigenspectra.
        The top row shows the true values for the map and the spectra, and
        the bottom row shows our inferred values.
        \label{fig:twospec}
    }
    \end{centering}
\end{figure}

\section{Discussion}
\label{sec:discussion}
%
\xxx{In progress...}
Evaluation of the model via Equation~(\ref{eq:deconv:F}) requires 
first the rotation of the $(\lmax + 1)^2$
$\kT$ kernels (each of length $W$) to the correct stellar 
rotational phase, an operation that scales approximately
as $W \lmax^3$,
followed by $(\lmax + 1)^2$
linear convolutions of the rotated $\kT$ kernels with the 
spectral components (each of length $K$), which scales
approximately as $W K \lmax^2$. Since typically $K \gg \lmax$,
the latter computation dominates and the entire evaluation 
takes $\mathcal{O}(W K \lmax^2)$ operations.

\xxx{We are not addressing differential rotation and convective blueshift.}
\xxx{Telluric lines could be treated much as in Figure~\ref{fig:twospec}.}
\xxx{Instrumental profile is another Toeplitz matrix.}

\section{Conclusions}
\label{sec:conclusions}
\xxx{Hello.}

>>>>

\begin{abstract} 
We derive a closed form, analytic solution to the problem of Doppler imaging
of stellar surfaces in the limit of negligible differential rotation and
convective blueshift. The model for the observed spectrum is linear in 
both the unbroadened spectrum and the surface map of the star, provided the
latter is expressed as a sum of spherical harmonics. As a result, the 
posterior distributions of both the unbroadened spectrum and the stellar map
conditioned on each other have a closed form, analytic solution that is 
computationally trivial to evaluate. When neither the unbroadened spectrum
nor the stellar map is known a priori, the global optimum can usually be
found via a fast iterative approach that also yields a lower bound on the
uncertainties.
\xxx{%
\textbf{Mention the following things:}
Advantages of data-driven modeling $\cdot$ %
Interpretable priors $\cdot$ %
Works on blended lines $\cdot$ %
Uses the whole spectrum $\cdot$ %
Yields uncertainties! $\cdot$ %
Probably works on slow rotators $\cdot$ %
Differentiable! $\cdot$ %
Fast! $\cdot$ %
Use in MCMC to marginalize over inclination and stuff
}
%
\href{https://github.com/rodluger/paparazzi}{\color{linkcolor}\faGithub}
\end{abstract}

%
\section{Introduction}
%
\xxx{%
\noindent Introduce the problem of Doppler imaging. Then cite the following papers:\\[0.5em]
\noindent\textbf{Misc.} \\
\noindent\citet{Luger2019} --- Surface modeling with spherical harmonics: \starry \\
\noindent\citet{Bedell2019} --- Data-driven spectral inference: \texttt{wobble} \\[0.5em]
%
\noindent\textbf{Theory} \\
\noindent\citet{Khokhlova1976,Goncharskii1977} --- Local line profile from observed spectrum\\
\noindent\citet{Vogt1983,Vogt1987} --- Seminal papers on Doppler imaging \\
\noindent\citet{Rice1989} ---  Another seminal paper\\
\noindent\citet{Deutsch1958,Deutsch1970,Falk1974} ---  Ylms + Doppler imaging\\
\noindent\citet{Unruh1995} --- sensitivity to line profile models\\
\noindent\citet{Piskunov1990} --- notes on regularization \\[0.5em]
%
\noindent\textbf{Application} \\
\noindent\citet{Crossfield2014} --- Global cloud map of Luhman 16B \\
\noindent\citet{Roettenbacher2016} --- Doppler imaging of $\zeta$ Andromedae \\
\noindent\citet{Roettenbacher2017} --- Doppler imaging of $\sigma$ Geminorum \\
}
%

@@

\subsection{The full solution}
\label{sec:full_solve}
In the previous two sections we showed how, if the stellar
rest-frame spectrum is spatially constant and known \emph{a priori} 
(from, say, a template spectrum), the posterior distribution
for the surface map $\bvec{y}$ is \emph{analytic} 
(Equations~\ref{eq:inverse:y1hat}
and \ref{eq:inverse:y1cov}). Conversely, if the surface map is known,
the posterior for the spectrum is also analytic 
(Equations~\ref{eq:inverse:shat} and \ref{eq:inverse:scov}). Both sets of
solutions require Gaussian priors on the quantities of interest, but these
are easily interpretable. A Gaussian prior on
the spherical harmonic coefficients of the surface map corresponds 
to the angular power spectrum of the surface features, whereas a Gaussian
prior on the spectrum can be used to correctly propagate uncertainties in
the stellar spectrum from observational uncertainties on the stellar
metallicity, temperature, etc.

@@

\subsection{Limb darkening}
\label{sec:ld}
Thus far we have avoided mention of limb darkening, whose effect
on the observed spectrum can often be significant. A proper treatment of
limb darkening requires solution of radiative transfer equations and
should account for the temperature and composition gradients in the
stellar atmosphere. However, in the spirit of data-driven inference, we 
can approximate the effect of limb darkening by weighting the spatial
component of our map by the limb darkening profile of the star, which can 
either be imposed or learned from the data. For a polynomial limb darkening
law of the form
%
\begin{align}
    \label{eq:ld:I}
    \frac{I(\mu)}{I(\mu = 1)} = 1 - \sum_{n=1}^{n_\mathrm{max}} u_n(1 - \mu)^n
    \quad,
\end{align}
%
where $\mu = z = \sqrt{1 - x^2 - y^2}$ is the radial coordinate on the 
projected disk and $u_n$ is a limb darkening coefficient, the effect of 
limb darkening on the stellar map can be expressed exactly as a linear 
operation on the spherical harmonic coefficient vector 
\citep{Luger2019,Agol2019}. This can be understood by noting that all terms
in Equation~(\ref{eq:ld:I}) are strictly polynomials in $x$, $y$, and $z$,
all of which can be expressed exactly as sums of spherical harmonics
\citep{Luger2019}. When weighting the surface intensity by the limb darkening
profile, the resulting intensity is simply a product of spherical harmonics,
which is itself a sum of spherical harmonics. Given a limb darkening law
of degree $n_\mathrm{max}$, we can always construct a matrix $\bvec{L}$
that transforms a spherical harmonic vector $\bvec{y}$ of degree
$l_\mathrm{max}$ to a limb-darkened spherical harmonic vector $\bvec{y'}$
of degree $l_\mathrm{max} + n_\mathrm{max}$. As an example, consider 
a map of degree $l_\mathrm{max} = 1$ and the linear limb darkening law
($n_\mathrm{max} = 1$).
The transformation matrix from $\bvec{y}$ to $\bvec{y'}$ is
%
\begin{align}
    \label{eq:ld:L}
    \setstackgap{L}{1.5\baselineskip}
    \bvec{L} =
        \frac{1}{1 - \frac{u_1}{3}}
        \begin{pmatrix}
        \quadquad1-u_1\quadquad & 0                         & \frac{u_1}{\sqrt{3}}      & 0                         \\
        0                       & \quadquad1-u_1\quadquad   & 0                         & 0                         \\
        \frac{u_1}{\sqrt{3}}    & 0                         & \quadquad1-u_1\quadquad   & 0                         \\
        0                       & 0                         & 0                         & \quadquad1-u_1\quadquad   \\
        0                       & 0                         & 0                         & 0                         \\
        0                       & \frac{u_1}{\sqrt{5}}      & 0                         & 0                         \\
        0                       & 0                         & \frac{2u_1}{\sqrt{15}}    & 0                         \\
        0                       & 0                         & 0                         & \frac{u_1}{\sqrt{5}}      \\
        0                       & 0                         & 0                         & 0          
        \end{pmatrix}
        \quad.
\end{align}
%
The columns of $\bvec{L}$ are constructed from the coefficient vectors of
each transformed spherical harmonic, which are in turn computed by 
multiplying each spherical harmonic by the spherical harmonic decomposition 
of the particular limb darkening law.
Figure~\ref{fig:ld} shows this transformation with $u_1 = 1$ applied to each
of the first four spherical harmonics.
%
\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=0.5\linewidth]{figures/ld.pdf}
    \oscaption{ld}{%
        Illustration of the limb darkening transformation given by
        Equation~(\ref{eq:ld:L}) for a linear limb darkening law with
        $u_1 = 1$. Each row shows a spherical harmonic before
        (left) and after (right) transformation by $\bvec{L}$.
        \label{fig:ld}
    }
    \end{centering}
\end{figure}
%

@@

The model remains linear in both the spherical harmonic coefficients and the
spectral components, and the procedures outlined in \S\ref{sec:inverse} may
still be followed to solve the inverse problem, provided we unroll these
matrices into vectors, letting
$\mathbf{s} = \mathrm{vec}(\boldsymbol{\mathbb{S}})$
and
$\mathbf{y} = \mathrm{vec}(\boldsymbol{\mathbb{Y}})$.
A detailed discussion of this general case is deferred to future work,
but we present an example of a two-component stellar spectrum in \S\ref{sec:eigen}
below.

@@

Given $\bvec{y_1}$ and $\bvec{s}$, we use Equations~(\ref{eq:deconv:F}) and~%
(\ref{eq:inverse:azero}) to
compute the data vector $\bvec{f}$
at $M = 16$ equally spaced rotational phases; these phases are shown in orthographic
projection (as an observer would see the star) in the center
panel of Figure~\ref{fig:spot_setup}. The synthetic spectra for each phase
are shown as the orange curves in the bottom panel of the figure (note the
different scaling, as all spectral lines become shallower due to the
broadening). We assume a wavelength-independent,
quadratic limb darkening law with $u_1 = 0.5$ and $u_2 = 0.25$. Finally,
we divide the flux by the baseline at each epoch (Equation~\ref{eq:norm:b})
and add random Gaussian noise with amplitude $\sigma_\bvec{f} = 10^{-4}$ to
simulate a very high signal-to-noise ratio (SNR) spectrum. The mean
standard deviation in each wavelength bin across all 16 epochs in the
noiseless data vector is about $10^{-2}$, so our choice of $\sigma_\bvec{f}$
corresponds to SNR $\sim$ 100 in each of the 201 wavelength bins of $\bvec{f}$.
%
%For reference, \citet{Vogt1987} report SNRs per 60 m\AA\, pixel, while our
%resolution elements are approximately $\lambda_r\Delta\lnlam/K \approx 100$ m\AA\,
%across, so our SNR in the same units as \citet{Vogt1987} is about 80.
%
% TODO: Perhaps Vogt computes the "signal" as the line depth, rather than
% as the amplitude of the perturbations? Check this.

@@

\subsection{Known baseline and spectrum}
\label{sec:spot_y1}
%
In our first experiment, we attempt to infer the stellar map assuming
we have perfect knowledge of the rest frame spectrum $\bvec{s}$ and
the baseline $\bvec{b}$ (Equation~\ref{eq:norm:b}). 
This is analogous to the case where a good physical model exists for
the stellar spectrum and simultaneous photometry is available to
constrain the baseline variations.
As we showed in
\S\ref{sec:solve_y}, the posterior distribution for the map coefficients 
$\bvec{y_1}$ is analytic in this case, with mean given by Equation~(\ref{eq:inverse:y1hat})
and covariance given by Equation~(\ref{eq:inverse:y1cov}). Note that
we must first scale the data vector $\bvec{f}$ by $\bvec{b}$ (which in this
case is known).

@@

\subsection{Known spectrum, unknown baseline}
\label{sec:spot_y1b}
%
We next investigate the case where the spectrum is known exactly, but
the baseline is not. This corresponds to the case discussed in 
\S\ref{sec:norm}. We use the solution to the linearized problem,
Equation~(\ref{eq:norm:yhat}), to obtain an initial guess for 
$\bvec{y_1}$. We then repeatedly solve
Equation~(\ref{eq:inverse:y1hat}) for $\bvec{y_1}$, given the baseline 
computed from the previous value of $\bvec{y_1}$ from Equation~(\ref{eq:norm:b}).
Because of the high SNR of our data, we find that it is necessary to employ
a simple tempering procedure to avoid getting stuck in local minima. We
introduce an artificial temperature $T$ that decreases every iteration
in equally spaced logarithmic steps $\Delta \log T$ from an initial temperature 
$T=T_0$ to $T=1$. This temperature is used to scale the data covariance
$\boldsymbol{\Sigma}_\bvec{f}$; in practice, we simply replace
$\boldsymbol{\Sigma}_\bvec{f}$ with $T \boldsymbol{\Sigma}_\bvec{f}$
in Equation~(\ref{eq:inverse:y1hat}). This has the effect of smoothing
out the likelihood space, reducing the likelihood that the solver
will get stuck in a local minimum. For this particular experiment, we
set $T_0 = 100$ and $\Delta \log T = -0.25$, for a total of $9$ steps of
the bilinear solver. 
%
In a final step, we run a gradient-descent optimizer, initializing it
at the solution obtained with the bilinear solver. We use
the \texttt{Keras} \citep{Keras} implementation of the
Nesterov momentum-enhanced Adam solver \citep[\texttt{NAdam};][]{NAdam},
running it for $N_\mathrm{nadam} = 80$ iterations at a learning rate
$L_\mathrm{nadam} = 10^{-4}$. Our loss function is simply the (negative)
log probability, equal to the sum of the log likelihood and log prior:
%
\begin{align}
    \label{eq:spotstar:logP}
    -\ln P &= 
    \frac{1}{2}
    \Big(
        \bvec{r}^\top \boldsymbol{\Sigma}_\bvec{f}^{-1} \bvec{r} \, + \,
        \bvec{y_1}^\top \boldsymbol{\Lambda}_\bvec{y_1}^{-1} \bvec{y_1} \, + \,
        (\bvec{b - 1})^\top \boldsymbol{\Lambda}_\bvec{b}^{-1} (\bvec{b - 1})
    \Big)
\end{align}
%
where $\bvec{r}$ is the residual (data minus model) vector and the last term
is a prior we include on the baseline $\bvec{b}$, which we force to be close
to a mean of unity with covariance $\boldsymbol{\Lambda}_\bvec{b} = (0.1)^2\bvec{I}$.

@@

Our results are shown in 
Figure~\ref{fig:spot_y1b}. The panels at the top are the same as those
in Figure~\ref{fig:spot_y1}, but now we also plot the inferred baseline
(lower left) and spherical harmonic coefficients (lower right) next to
the true values. We recover the \texttt{SPOT} map just as well as before,
except for the presence of a strong dipolar artifact; the background
intensity level is higher in the northern hemisphere and lower in the
southern hemisphere. This effect is closely related to the fact that
we do not know the true baseline. While we are able to infer the
shape of the baseline exceedingly well (lower left panel), there is a large
offset between the true value and our inferred value. This is due to
a fundamental degeneracy involving the 
$Y_{1,-1}$ (\raisebox{-0.2em}{\includegraphics[width=1em]{figures/Y1m1.pdf}})
spherical harmonic coefficient, which, because a large portion of the southern
hemisphere is never in view, cannot be constrained by the data. The 
magnitude of the true value of the coefficient is quite large (lower right
panel), since the northern hemisphere \emph{is} significantly darker than
the southern hemisphere; however, the data do not constrain this, and 
our prior shrinks all coefficients toward zero. The net effect is that the
region surrounding the spot in the northern hemisphere gets brighter to 
even out the difference between the two hemispheres. This causes the 
visible portion of the star to get brighter, raising the baseline
relative to the true value.

Note, importantly, that even in cases where the baseline is known (from, say,
simultaneous photometry), it is typically only known in a \emph{relative} 
sense. This degeneracy will typically apply in that case as well, so in
principle the dipole term will always be unconstrained for stars at
similar inclinations. For inclinations closer to $90^\circ$ this is much 
less of an issue, although (as we will see below) other degeneracies show up.

Finally, we note that the pixelwise uncertainty shown is a strict
\emph{lower bound} on the map uncertainty, since it is the uncertainty
conditioned on the inferred value of the baseline. Sampling methods
such as MCMC are required to obtain the \emph{true} uncertainty on the 
map. We discuss this further in \S\ref{sec:discussion}.

\subsection{Unknown spectrum, unknown baseline}
\label{sec:spot_y1bs}
%
Suppose we are handed a series of spectra in a wavelength regime where
we do not have a good physical model for the line generation mechanisms,
or a wavelength regime with many line blends for which we do not have
reliable templates. Or perhaps the spectra are of a star whose
rotational broadening does not dominate over other broadening mechanisms,
in which case our template might not be accurate enough to allow us to
tease out the small Doppler signals. In these cases, we may want to relax
the assumption that we know the rest frame spectrum $\bvec{s}$ exactly,
and instead try to \emph{learn} it from the data.

Given the same data as before, we now try to learn both the stellar map
and the stellar spectrum, conditioned only on the wide Gaussian priors
discussed above. We must follow the iterative procedure presented in
\S\ref{sec:full_solve}, adjusting for the fact that we do not know
the true baseline, either (\S\ref{sec:norm}).

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=0.49\linewidth]{figures/spot_y1bs_timeseries.pdf}
    \includegraphics[width=0.49\linewidth]{figures/spot_y1bs_rect.pdf}
    \\[0.5em]
    \includegraphics[width=0.405\linewidth]{figures/spot_y1bs_baseline.pdf}
    \includegraphics[width=0.575\linewidth]{figures/spot_y1bs_coeffs.pdf}
    \\[0.5em]
    \includegraphics[width=0.98\linewidth]{figures/spot_y1bs_spectrum.pdf}
    \oscaption{spot_y1bs}{%
        Similar to Figure~\ref{fig:spot_y1}, but assuming we have no
        information about either the spectrum, the baseline, or the map
        coefficients. The bottommost panel now shows the true rest frame
        spectrum (blue line), the initial guess based on a deconvolution
        (dashed orange line), and the inferred spectrum (orange line, with
        shaded $1\sigma$ uncertainty).
        \label{fig:spot_y1bs}
    }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
    \includegraphics[width=0.49\linewidth]{figures/spot_y1bs_low_snr_timeseries.pdf}
    \includegraphics[width=0.49\linewidth]{figures/spot_y1bs_low_snr_rect.pdf}
    \\[0.5em]
    \includegraphics[width=0.405\linewidth]{figures/spot_y1bs_low_snr_baseline.pdf}
    \includegraphics[width=0.575\linewidth]{figures/spot_y1bs_low_snr_coeffs.pdf}
    \\[0.5em]
    \includegraphics[width=0.98\linewidth]{figures/spot_y1bs_low_snr_spectrum.pdf}
    \oscaption{spot_y1bs_low_snr}{%
        Same as Figure~\ref{fig:spot_y1bs}, but for $10\times$ lower SNR.
        We still recover the \spot map and the input spectrum, albeit with
        higher uncertainty, as expected.
        \label{fig:spot_y1bs_low_snr}
    }
    \end{centering}
\end{figure}

Because the problem is generally not convex, we need a good starting
guess for either the spectrum or the map coefficients. We note that
we can very crudely approximate the rest frame spectrum by deconvolving the 
average observed spectrum with the first term of the convolution kernel
$\kT$ (\raisebox{-0.1em}{\includegraphics[width=2em]{figures/kT00.pdf}}).
This is equivalent to assuming the average map of the star
across all observations is featureless, and the average spectrum is equal
to the (symmetrically) rotationally-broadened version of the rest frame
spectrum. In practice, this is a fairly poor approximation, but we find that it
suffices as an initial guess. Given this starting value of $\bvec{s}$, we 
compute an initial guess for $\bvec{y_1}$ using the linear baseline
approximation. We then iterate between solving for $\bvec{s}$ and
solving for $\bvec{y_1}$, always using the previous value of the 
baseline to compute $\bvec{y_1}$. As in the previous section, we find it
necessary to implement a tempering scheme, except with a much higher
starting temperature $T_0 = 5,000$ and a slower decay $\Delta\log T = -0.04$.
We again refine our solution with \texttt{NAdam}, this time setting
$N_\mathrm{nadam} = 3,000$ and $L_\mathrm{nadam} = 2.5\times 10^{-5}$. Since
we are now solving for the spectrum as well, we include the negative log prior 
term for the spectrum, 
$\frac{1}{2}(\bvec{s - 1})^\top \boldsymbol{\Lambda}_\bvec{s}^{-1}(\bvec{s - 1})$, 
in Equation~(\ref{eq:spotstar:logP}).

Figure~\ref{fig:spot_y1bs} shows the results. The top portion of the figure is
the same as before. A panel at the bottom has been added showing the true
rest frame spectrum (blue), the initial guess based on the deconvolution
(dashed orange), and the inferred spectrum (orange). We recover the \spot
feature as before, albeit with more noise, particularly south of about
$+30^\circ$ latitude. Our uncertainties are somewhat underestimated, as 
expected, since they are computed conditioned on the optimum value of 
the spectrum. On the other hand, we recover the spectrum exceedingly well,
with low uncertainty and virtually no bias anywhere.

To test our technique on more realistic data, in Figure~\ref{fig:spot_y1bs_low_snr}
we repeat the experiment on data with $10$ times higher uncertainty,
setting $\sigma_\bvec{f} = 10^{-3}$. Because the higher uncertainty increases
the volume of the region surrounding the global optimum, it is
actually significantly \emph{easier} for the algorithm to converge on the
best solution, so we set $T_0 = 500$, $\Delta\log T = -0.05$, 
$N_\mathrm{nadam} = 500$, and $R_\mathrm{nadam} = 10^{-4}$. Interestingly,
while the uncertainty on the spectrum increased, the quality of the
recovered \spot map did not noticeably degrade. This suggests we did not
quite find the global optimum in Figure~\ref{fig:spot_y1bs}. We discuss
this further in \S\ref{sec:discussion}.

\subsection{Spatially-variable spectrum}
\label{sec:twospec}
%
In our final experiment, we relax the assumption that the stellar
rest frame spectrum is spatially constant, allowing for different
spectral features inside and outside of the \spot. This is the case
for, say, chemically peculiar Ap stars, whose abundance of certain
elements is significantly different within spotted regions on their
surfaces. It is also the case in general for spots that are significantly
cooler than the rest of the photosphere, where the relative abundance
of different ionization states or molecular features may be different
than in the hotter, unspotted regions.

The problem setup is the same as that discussed in \S\ref{sec:eigen}, where
our full map vector $\bvec{a_0}$ is the outer product of a spectral matrix
$\mathbb{S}$, whose $J$ columns are the different spectral components, and
a spherical harmonic coefficient matrix $\mathbb{Y}^\top$, whose $J$ rows
are the corresponding coefficient vectors. For simplicity, in this experiment
we consider only $J = 2$ spectral components and further require that the
first component is spatially constant, i.e., the first row of 
$\mathbb{Y}^\top$ is the vector $(1 \quadquad \bvec{0}^\top)$. The second
row of $\mathbb{Y}^\top$ is the vector $(1 \quadquad \bvec{y_1}^\top)$, where
$\bvec{y_1}$ is the (unknown) spherical harmonic expansion of the spot. The
two columns of $\mathbb{S}$ are $\bvec{s_1}$ and $\bvec{s_2}$, the unknown
spectral components.

\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/twospec.pdf}
    \oscaption{twospec}{%
        An example of a star whose spectrum is different within the spot
        (\texttt{B}) and outside of the spot (\texttt{A}). Everywhere on 
        the surface, the spectrum is a linear combination of two
        ``eigenspectra'' ($s_1$ and $s_2$).
        As in Figure~\ref{fig:spot_y1bs}, we assume
        we do not know the map, the baseline, or either of the eigenspectra.
        The top row shows the true values for the map and the spectra, and
        the bottom row shows our inferred values.
        \label{fig:twospec}
    }
    \end{centering}
\end{figure}
%
\xxx{Figure~\ref{fig:twospec} shows...}

\section{Discussion}
\label{sec:discussion}
%
\subsection{Scaling of the computations}
\label{sec:scaling}
%
Evaluation of the model via Equation~(\ref{eq:deconv:F}) requires 
first the rotation of the $(\lmax + 1)^2$
$\kT$ kernels (each of length $W$) to the correct stellar 
rotational phase, an operation that scales approximately
as $W \lmax^3$,
followed by $(\lmax + 1)^2$
linear convolutions of the rotated $\kT$ kernels with the 
spectral components (each of length $K$), which scales
approximately as $W K \lmax^2$. Since typically $K \gg \lmax$,
the latter computation dominates and the entire evaluation 
takes $\mathcal{O}(W K \lmax^2)$ operations. \xxx{More on this.}
